Hierarchical ARIMA

### 2020-08-16 KB_modeling.Rmd
Combined new playoff data and added ar terms

dat <- dat %>%
  mutate_at(vars(DivWin, WCWin, LgWin, WSWin), funs(as.logical)) %>%
  mutate(
    playoffs = case_when(
      WSWin ~ "Won WS",
      LgWin ~ "Won League",
      WCWin  ~ "Won Conference",
      DivWin  ~ "Won Division",
      TRUE ~ "Did not qualify"
    ),
    playoffs = forcats::fct_inorder(playoffs),
    phase = case_when(
      month < 04 | month > 10 ~ "Summer",
      month == 04 ~ "Opening",
      month < 10 & month > 04 ~ "Season",
      month == 10 ~ "Playoffs"
    ),
    phase = forcats::fct_reorder(phase, month),
    is_playoffs = as.integer(phase == "Playoffs")
  ) %>%
  mutate_at(vars(month, year, team), funs(factor))

summary(dat$playoffs)

## playoffs variable
#  Did Not make Playoffs or WC Play-in 
#  Played in WC Play-In
#  Eliminated Round 1
#  Eliminated Round 2
#  Lost WS
#  Won WS
```

@@

annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll, data = by_year) ## this should be an AR model

@@

Much more evened out for Dodgers and Dbacks.  Why not for Cubs?  Probably because
of the later increasingtrend they have.

@@

```{r}
final_model <- 
  dat  %>%
  lm(ghits_e2 ~ phase + is_playoffs:playoffs, data = .) ## needs to be AR model instead

>>>>

## Added new dataset for playoff
# broken up into 6 different factors
# - Won WS (1 team)
# - Lost WS (1 team)
# - Lost Semi (2 teams)
# - Lost Quarter (4 teams)
# - Lost WC (2 teams)    NOTE: wildcard game started in 2012 so this variable is only from 2012-2019.
# - No Playoffs (22 teams from 2004-2011 and 20 teams from 2012-2019)

playoff_success <- read.csv("MLB_Postseason_Success.csv", stringsAsFactors = FALSE) %>%
  mutate_all(na_if,"") %>%
  replace(is.na(.), "No Playoffs")

#combining new palyoff dataset with dat
dat <- left_join(dat, playoff_success, by = c("team", "year")) %>%
  select(-c(DivWin, WCWin, LgWin, WSWin)) # deselecting unnecessary variables

dat <- dat %>%
  # mutate_at(vars(DivWin, WCWin, LgWin, WSWin), funs(as.logical)) %>%
  mutate(
  #   playoffs = case_when(
  #     WSWin ~ "Won WS",
  #     LgWin ~ "Lost WS",
  #     WCWin  ~ "Won Wild Card",
  #     DivWin  ~ "Won Division",
  #     TRUE ~ "No Playoffs"
  #   ),
    playoffs = forcats::fct_inorder(playoffs), # not exactly sure what this line of code does? do we still need?
    phase = case_when(
      month < 04 | month > 10 ~ "Summer",
      month == 04 ~ "Opening",
      month < 10 & month > 04 ~ "Season",
      month == 10 ~ "Playoffs"
    ),
    phase = forcats::fct_reorder(phase, month), # not exactly sure what this line of code does? do we still need?
    is_playoffs = as.integer(phase == "Playoffs")
  ) %>%
  mutate_at(vars(month, year, team), funs(factor))

summary(dat$playoffs)
```

@@

# NOTE to self: check this again to see if i did it right

annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll + lag(ghits_e1, 1), data = by_year) ## this should be an AR model, annual

# i used the lag function , but examples I see have used dynlm package

@@

Much more evened out for Dodgers and Dbacks.  Why not for Cubs?  Probably because
of the later increasing trend they have.

@@

```{r}
final_model <- 
  dat  %>%
  lm(ghits_e2 ~ phase + is_playoffs:playoffs + lag(ghits_e2, 1), data = .) ## needs to be AR model instead, monthly
